\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-06-01 eg6-letters/NIM/EG6_RTPC_NIM.tex

The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.
Figure \ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width.  The shift between the rows allows to reduce aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in figure~\ref{fig:PADs} 
shows how pads are grouped to 16 channels pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC \cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and driving the 6m long ribbon 
cable that connects to it.

@@

The trigger signal also initiates RCU readout operation from FEC boards. All 
the measured samples from active channels are reported, together with a channel 
identifier and a time-stamp, to the ROC application, that in turns sends them 
on the main event builder. During data reconstruction, the acquired samples are 
processed to obtain, for each readout pad, the accumulated charge (ADC) and the 
pulse time (T). Since pulse time was obtained as the time-stamp of the first 
sample above threshold, referred to to the trigger time, the resolution is 
equivalent to the ALTRO sampling time, 100~ns.

@@

Two quantities are recorded at the readout board, time (T) and signal amplitude 
(ADC). The timing information is used to infer the origin of the charge and 
then the trajectory of the detected particle
resulting in momentum measurement. Going from a collection of 
times to a momentum measurement requires a good knowledge of the drift speed 
and drift paths followed by the electrons released in the gas. The recorded ADCs give 
the deposited energy per unit of length ($\small{\frac{dE}{dX}}$) which, 
together with the momentum calculated from the trajectory, enables particle 
identification.

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We still always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the asimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the one obtained while performing the last iteration of this long 
calibration process.

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but cylindrical symmetry insures they always travel the same 
distance. By identifying the maximum time T measured, we can infer the drift 
speed of the electrons in the RTPC.\\

>>>>

The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.
Figure \ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width.  The shift between the rows allows to reduce aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in figure~\ref{fig:PADs} 
shows how pads are grouped to 16 channels pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC \cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and driving the 6~m long ribbon 
cable that connects to it.

@@

The trigger signal also initiates RCU readout operation from FEC boards. All 
the measured samples from active channels are reported, together with a channel 
identifier and a time-stamp, to the ROC application, that in turns sends them 
on the main event builder. During data reconstruction, the acquired samples are 
processed to obtain, for each readout pad, the accumulated charge (ADC) and the 
pulse time (T). Since pulse time was obtained as the time-stamp of the first 
sample above threshold, the resolution is equivalent to the ALTRO sampling 
time, 100~ns.

@@

The timing information is used to infer the origin of the charge and 
then the trajectory of the detected particle
resulting in momentum measurement. Going from a collection of 
times to a momentum measurement requires a good knowledge of the drift speed 
and drift paths followed by the electrons released in the gas. The recorded ADCs give 
the deposited energy per unit of length ($\small{\frac{dE}{dX}}$) which, 
together with the momentum calculated from the trajectory, enables particle 
identification.

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the asimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the one obtained while performing the last iteration of this long 
calibration process.

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but cylindrical symmetry insures they always travel the same 
distance. By identifying the maximum time measured ($T_{Max}$), we can infer the drift 
speed of the electrons in the RTPC.\\
